Cache Oblivious B-trees

### 2011-12-09 paper.tex
Adding the abstract

This paper presents a survey of cache-oblivious B-tree data structures. Memory
hierarchies . Since memory transfers often dominate the runtimes of an
algorithm it makes sense to develop data structure that are optimized not in
terms of the number of instructions it executes; but rather, by the number of
memory accesses and transfers it has to perform. Among these so called
``external-memory'' data structures, the B-tree is widely used as it is able
to perform \Search in an optimal number of memory transfers. However, a short
coming of B-trees is that they require knowledge of the size of the memory
blocks, thus it is a cache aware data structure. This gave rise to the notion
of cache-oblivious algorithms, e.g. data structures that perform well
regardless of the size of the memory blocks or the number hierarchies. Bender
et al. introduced the cache-oblivious B-tree which ...

>>>>

This paper present a survey on cache-oblivious B-trees (CO B-trees). The CO B-tree
was first introduce by Bender et al in 2000. It is an external memory data structure meaning
that it optimizes the number of memory transfers if performs rather than the number of
CPU instructions that it executions. Unlike other external memory data structure, like
the widely use B-tree, the CO B-tree does not require any information about the
number of level in the machine's memory hierarchy, the block size and capacity at each level, or the 
relative speeds for memory transfer. Hence, the CO B-tree is ``cache-oblivious''.
Despite this, the CO B-tree is able to achieve the same
optimal search bound as the B-tree and near optimal insertion and deletion bounds as well.
In this paper we first describe the original CO B-tree and explain how it is able to achieve 
its memory performance. We then present and analyze three followup CO B-tree data structures
 that improved upon shortcoming of the original: the cache-oblivious string B-tree, which introduces support 
for variable length keys, the cache-oblivious streaming B-tree, which incorporates improvements to the
insertion bound at the expense of the search bound, and lastly, the cache-oblivious concurrent B-tree, which 
introduces support for concurrency.
